--- 
title: "PSTAT 194 Final Project" 
author: "Ayden Kelleher, Owen Feng, Yuxin Chen, Henry Bergeson" 
date: "2026-03-04" 
output: pdf_document
--- 

```{r setup, include=FALSE} 
  knitr::opts_chunk$set(echo = TRUE) 
``` 

# Introduction: 

For our network analysis, we chose for our network to be the Multi-scale Attributed Node Embedding Network Wikipedia page-page networks on Squirrels. [This network](https://snap.stanford.edu/data/wikipedia-article-networks.html) consists of a subset of Wikipedia articles, where the nodes represent wikipedia articles on squirrels and the edges represent mutual hyperlinks between those articles. Because edges only exist when two articles link to each other, the graph is undirected. 

Using the `summary` function, we can see that our network has 5201 nodes, 217073 edges, that our graph is unweighted and undirected, and that it contains two node attributes. The first attribute, `target`, is the average monthly traffic to each article from October 2017 - November 2018. The second attribute, `features`, is a list containing a collection of numbers corresponding to specific nouns appearing in the articles. However, we were unable to determine which nouns corresponded to which number. Histograms of these attributes are shown [here](#snippet-2) to provide a basic overview of their distributions. 

The objective our analysis is to explore the structural properties of the squirrels Wikipedia article network and to investigate whether certain attributes influence the amount of traffic a node receives. 

# Methodology: 

## Tools and Packages Used:

We first loaded several `R` packages required for the analysis, including `igraph`, `tidygraph`, `jsonlite`, and `tidyr`, along with the dataset shown in [Code Snippet 1](#snippet-1). 

## Steps followed in data processing. 

The graph data was provided as an edge list. We first converted the edge list into a matrix and then turned the matrix into a graph using the `graph_from_edgelist()` function from the `igraph` package. The feature data was stored as JSON files and was loaded using the `fromJSON()` function from the `jsonlite` package. Lastly, the target data, which was the average monthly traffic for each article, was loaded in using the `read.csv()` function from base `R`. These datasets were then added to the graph as vertex attributes. 

# Generate Induced Subgraphs: 

## Create four induced subgraphs with 200 random nodes each from your dataset. 

To create our induced subgraphs, we first attempted to [sample 200](#first-attempt-at-creating-a-subgraph) with 200 random nodes from our graph. However, this resulted in many "isolated" nodes: 79 out of 200 sampled nodes had degree zero,meaning they were not connected to any other nodes. This was unexpected, as the original graph $g$ contains no nodes with degree zero. As a result, purely random sampling caused much of the network’s structure to disappear. 

To resolve this issue, we instead drew a random sample of 40 nodes from the network. Then, for each of these nodes, we sampled up to five of their neighbors, or all of the neighbors if the node had fewer than five. From this resulting set of nodes, we selected 200 nodes to form our induced subgraph. Although this method introduces some bias, it preserves much of the network's connectivity and structure than purely random sampling. We used this approach to generate each of our four subgraphs, which can be seen in [subgraph 1](#subgraph-1), [subgraph 2](#subgraph-2), [subgraph 3](#subgraph-3), and [subgraph 4](#subgraph-4). More detailed visualizations of these networks are provided later. 

# Visualize the Networks: 

## Use three different layouts from the igraph package to visualize each network. 

For our visualizations, we chose 3 different layouts: Fruchterman-Reingold, Kamada-Kawai, and Nicely. The plots for each of these layouts can be found [here](#three-different-visualizations). 

## Experiment with node and edge attributes to enhance visualizations.

To enhance out visualizations and make them more interpretable, we chose to do a few different things: 

- Adjusted the node size by the following function of degree `3 + 8 * (degree(simp_graph) / max(degree(simp_graph)))`.

- Adjusted the node size by the attributes, specifically log internet traffic and log length of the wordlist. - Colored in nodes based on Louvain communities to make the groupings more obvious.

- Removed the node labels. Our nodes are labelled 0 to 5200. This gives us no information about the corresponding articles, so not showing the degree label doesn't lose any information. Even if we had the article names, it would be very difficult to read. 

- Made the lines between the nodes more see through so they take up less of the visual space. 

# Analyze Network Metrics: 

## Explore various network, node, and edge metrics. Interpret these metrics in the context of your network. 

### Degree 

The degree of a node measures the number of edges connected to that node. In directed graphs, this can be separated into in-degree and out-degree, but since our graph is undirected, the degree simply counts the total number of edges connected to each node. We first examined the [histogram of node degrees](#analyzing-network-metrics-degree) for the entire network. From our plots, we can see that the distribution is heavily right-skewed, with a large concentration of nodes in the $[0,100]$. This seems to indicate that most Wikipedia articles in the network link to relatively few other articles, while a smaller number of nodes have much higher connectivity. This distributions suggests that we have a few highly connected articles that act as hubs within our network. 

When we look at individual subgraphs, we observe a similar pattern. Each subgraph's degree distribution is right-skewed, meaning that most nodes have relatively low degree while a small number of nodes are highly connected. The most connected nodes in the subgraph (highlighted in yellow) have degrees just under 100, which is a slightly less than half of the nodes in each subgraph. This behavior is consistent with our full network, where the most connected nodes are also linked to roughly half of the other nodes. An interesting pattern that appears in each subgraph is bimodality, with a secondary peak in the degree range $[30, 50]$. 

It is unclear whether this second peak is a result of a true pattern in the data or is an artifact from the bias introduced by our sampling method. Recall that we first randomly sampled 40 nodes, and then sampled neighbors of those nodes to construct the subgraph. Because nodes with higher degrees are connected to more nodes, they are more likely to appear when sampling neighbors of those nodes. As a result, this secondary peak may be partially explained by the way we sampled our data, which increases the likelihood that higher-degree nodes are included in the subgraphs. 

### Closeness 

The closeness metric is a measure of how close every other node is to a specific node: 

$$C_C(v_i) = \frac{n-1}{\sum_{i\not=j}d(v_i,v_j)}$$ 

It is essentially the inverse of the sum of the distances from node $v_i$ to all other nodes in the network. The value is normalized so if the distance for all nodes was 1, then it would have a closeness of 1. This metric describes how centrally located an article is based on how quickly it can reach the rest of the network through chains of hyperlinks. An article with high closeness centrality is, on average, fewer link steps away from other articles, meaning it is more efficiently connected within the network.

Looking at the [closeness centrality plots](#analyzing-network-metrics-closeness), we can see that although nodes near the edges of the graph appear smaller, most other nodes are similar in size. This indicates that most nodes have similar closeness values, suggesting they are at roughly similar distances from the rest of the network.

### Betweenness 

The betweenness centrality is a measure of how often a node lies on the shortest paths between other pairs of nodes: 

$$C_B(v_i) = \sum_{i\not=j\not=k} \frac{\sigma_{jk}(v_i)}{\sigma_{jk}}$$ 

Where $\sigma_{jk}$ is the number of shortest paths between $v_j$ and $v_k$ and $\sigma_{jk}(v_i)$ is the number of shortest paths that pass throguh v_i. The resulting proportion describes how often the shortest path passes through node $v_i$. 

Looking at the [betweenness centrality plots](#analyzing-network-metrics-betweenness), we can that the histogram is skewed right, indicating that most nodes have relatively low betweenness centrality while only a small number of nodes have high betweenness values. When examining the network visualization with vertex size as log betweenness centrality, we see variation in node sizes. Nodes near the edges of the network are often extremely small, while nodes near the center tend to be larger, which is expected since central nodes are more likely to lie on shortest paths between many pairs of nodes. However, we also observe some exceptions, like larger nodes near the edges or smaller nodes closer to the center. This suggests that certain nodes located near the center of the graph are not critical for connecting different parts of the network, since many shortest paths between nodes can bypass them. Larger nodes near the edges may represent a bridge between an adjacent cluster and the rest of the rest of the network, causing many shortest paths between those regions to pass through that node. 

### Eigen Metric

The eigenvector centrality metric measures the importance of a node based on both how many nodes it's connected to and how important those nodes are. As described in the documentation, "eigenvector centrality scores correspond to the values of the principal eigenvector of the graph's adjacency matrix; these scores may, in turn, be interpreted as arising from a reciprocal process in which the centrality of each actor is proportional to the sum of the centralities of those actors to whom he or she is connected. In general, vertices with high eigenvector centralities are those which are connected to many other vertices which are, in turn, connected to many others (and so on)" ([igraph developers, 2024](https://r.igraph.org/reference/eigen_centrality.html)). In other words, a node receives a higher score, and is therefore considered more important, if it is connected to other nodes that are themselves well connected.

Looking at the plots of subgraphs [here](#analyzing-network-metrics-eigencentrality), we can see that eigen centrality is distributed on $[0,1]$, where 1 is more connected. Looking at our subgraphs, we see the mode is in the score range $[0,.1]$, with a second, much smaller, peak at about $[.4,.5]$. Comparing these distributions to our degree distributions, we observe a similar shape: a large concentration near zero and a smaller peak toward the middle of the range. This similarity is expected because the two metrics are related. Nodes with higher degree are more likely to be connected to other well-connected nodes, which increases their eigenvector centrality. 

When visualizing the entire network, the raw eigen metric is difficult to see because most nodes appear extremely small. To make our plots more interpretable, we applied a transformation to the eigenvector centrality values and used them to scale our verticies. To give an idea of what this transformation does, we also plotted the histogram of the transformed data for the first subgraph. 

Looking at the network plots, we can see that many of the central nodes are grouped together, while several of the less connected peripheral nodes appear much smaller, a pattern that seems to be most notable in subgraph 3. This pattern reveals that highly central nodes tend to be connected to other highly connected nodes, while peripheral nodes have fewer or weaker connections. This pattern shows eigen centrality is a good measure of centrality. 

# Community Detection: 

## Run community detection algorithms 

We chose to run [three different community detection algorithms](#we-choose-fast-greedy-louvain-and-walktrap-top-three-methods-on-our-lecture-slide-and-covered-in-ica-to-perform-our-community-detection.), Fast-Greedy, Louvain, and Walktrap. 

The Fast-Greedy community detection algorithm works by first declaring each node as its own community and then iteratively merging communities in a greedy manner, choosing the merge that produces the largest increase in modularity at each step. One drawback of this method is that, in larger networks, it may fail to detect smaller communities. 

The Louvain community detection algorithm works by first assigning each node to its own community and then iteratively moving nodes into neighboring communities if it increase the modularity of the network. After no further improvements can be made, the detected communities are collapsed into super-nodes, and the process repeats on this reduced network. 

The Walktrap community detection algorithm works by performing short random walks on the network. The idea is that nodes within the same community are more likely to be visited together during these walks, since random walks tend to remain within densely connected parts of the graph. 

## Analyze and interpret the resulting communities 

Comparing the results from the algorithms was interesting. Starting with the [modularity of our subgraphs](#and-then-we-compare-modularity), we see that none of the graphs have particularly high modularity. The results are shown below: 

$$
\begin{array}{rccc}
\text{subgraph} & \text{Fast-Greedy} & \text{Louvain} & \text{Walktrap} \\
1 & 0.1911257 & 0.2461304 & 0.15813260 \\
2 & 0.3522581 & 0.3608913 & 0.29551518 \\
3 & 0.1545764 & 0.2025717 & 0.10072104 \\
4 & 0.2772735 & 0.3141144 & 0.08605815
\end{array}
$$

From lecture, we learned that modularity over $0.3$ suggests meaningful community structure. However, only subgraphs 2 and 4 hadimitation any communities with modularity at least this high. Subgraphs 1 and 3 had modularity of $0.246$ and $0.202$ for their most modular community structrues, which aren't much smaller than $0.3$. This suggests that our network has a weak community structure. 

### Monte Carlo Estimate of Modularity 

To test if our community structure was meaningful, we ran a [monte carlo](#monte-carlo-modularity) to estimate the modularity of random networks with the same number of nodes and edges. We generated many graphs with the same number of vertices and edges, where the vertices had approximately the same degree distribution as the original network but the connections between nodes were assigned randomly. Comparing the results, we found that the observed modularity was much larger than the modularity of the simulated graphs, suggesting that the community structure in our network is not due to random chance and is therefore meaningful.

### Analyzing communities

Looking at the number of individual communities for each subgraph, we can see the following: 

$$
\begin{array}{rccc}
\text{subgraph} & \text{Fast-Greedy} & \text{Louvain} & \text{Walktrap} \\
1 &6 &5 &31 \\ 
2 &6 &6 &27 \\ 
3 &4 &7 &27 \\
4 &6 &5 &20 
\end{array}
$$

This gives us the biggest differences between the different community finding algorithms. We can see that both the Fast-Greedy and the Louvain algorithms ended up with a small amount of communities, around 6 for each subgraph. For the Walktrap, each subgraph had many more communities, with 31, 27, 27, and 20 for each of the subgraphs respectively. Looking at [the plots](#and-we-visualize-with-community-coloring-short-eg) of the subgraphs with the community coloring, we can see that the Walktrap has many isolated nodes that count as their own communities. This tells us if a node is relatively disconnected from the rest of the network, the Walktrap will have a hard time placing it in a larger community. 

# Adjacency Matrix: 

## Examine the adjacency matrix of your network.

We felt it was important to show at the unordered adjacency matrix to illustrate what an uncorrelated adjacency matrix looks like. Looking at the [unordered adjacency matrix](#unordered-adjacency-matrix) for subgraph 1 using the Louvain method, we observe very little visible structure. The only noticeable pattern are a few horizontal and vertical lines, which represent nodes with many connections, and a few faint blocks, indicating groups of nodes connected to similar sets of nodes. This lack of structure is expected since our nodes have not yet been reordered based on their communities. 

## Reorder vertices to highlight patterns. 

After [reordering our nodes](#ordered-adjacency-matrix), we can see a clearer pattern. There are five squares of varying density along the diagonal, which correspond to groups of nodes that are more densely connected with one another. This matches the five communities identified in Subgraph 1 by the Louvain method, confirming that these blocks represent the detected communities. Some observations we made were that There are five squares of varying density along the diagonal. This matches the five communities identifiedin subgraph 1 by the Louvain method, confirming that these blocks represent our detected communities. Some observations we made were that: 

- The first square is much more densely filled than the others, indicating that the nodes within this community are more strongly connected. 

- We can see many horizontal/vertical lines, particularly in the region corresponding to nodes in the first community. These lines represent nodes with many connections, meaning that this community has a higher average degree than the others. 

- The other four communities seem have a much lower degree on average, with noticeably more empty space within their blocks. This indicates that the graph is much less connected among themselves, highlighting that the subgraph overall is sparse. 

Because our subgraph consists of a single connected component, we cannot reorder our network by connected components. 

One question we considered is whether our sampling strategy affected this result. Recall that we first selected 40 random nodes, then sampled up to five neighbors of each node (or all neighbors if the node had fewer than five). Then, from this pool, we selected 200 nodes to construct the subgraph. Because many of the sampled nodes were chosen as neighbors of other nodes, the selection process was not entirely random but instead biased toward higher-degree nodes, since nodes with more connections are more likely to appear as neighbors. As a result, the nodes in our subgraphs are likely to be more connected than a typical node in the full network. This bias may therefore lead to fewer but larger communities than expected in our subgraphs. 

# Conclusion:

## Summary of insights gained. 

Looking at our network, we can see that there is significant structure to how links form between Wikipedia articles. From the plots of the subgraphs, we can see there are both centrailized and more peripheral nodes. We did not find a strong relationship between Internet traffic and node centrality, or between length of the word list and centrality. In other words, node size appeared to be distributed independently of whether a node was located near the center or the edge of the plot. However, we suspect that a relationship may become clearer if we were able to analyze larger subgraphs. 

While our graph had meaningful community structure, there were also many connectionss between communities, resulting in a lower modularity than was obsered in other networks. This makes intuitive sense in the context of Wikipedia: individual pages often contain many hyperlinks, some pointing to closely related topics and others to pages that are only loosely related. These numerous links naturally create connections between communities, reducing the overall modularity of the network. 

## Possible future directions for further analysis. 

- Analyze larger subgraphs, or maybe even the entire netowrk. 

- Closer look at attributes: a.) Ideally identify which words correspond to which `features` indices in the dataset. b.) If possible, analyze what percentage of internet traffic comes from articles within this graph compared to outside sources. This may provide more insight into how traffic relates to the network structure. 

# References: 
Rozemberczki, Benedek, Carl Allen, and Rik Sarkar. “Multi-Scale Attributed Node Embedding.” Journal of Complex Networks, vol. 9, no. 2, 2021. https://arxiv.org/abs/1909.13021. 

# Appendix

## Snippet 1

[back to report](#load-and-prepare-the-data)

```{r}
#install.packages('igraph')
library(igraph)
#install.packages('tidygraph')
library(tidygraph)
library(jsonlite )
library(tidyr)

edges_matrix <- as.matrix(read.csv("musae_squirrel_edges.csv"))

json_data <- fromJSON("squirrel.json")

target_data <- read.csv('squirrel_target.csv')

edges_matrix[,1] <- factor(edges_matrix[,1]) # if we dont change this to factor, then there are too many id names idk. If you don't believe me comment this out :/

g <- graph_from_edgelist(edges_matrix, directed = FALSE)

V(g)$features <- json_data

V(g)$target <- target_data$target

```

```{r}
length(V(g)) #this is exactly what the help page says
length(E(g)) #this is quite a few more edges than the help page says
```

## Snippet 2

[back to report](#identify-basic-network-elements-vertices-edges-weights-directionality-etc)

```{r}
length(V(g))
length(E(g))
is_weighted(g)
is_directed(g)
summary(g)
```

A brief look at the network features. These are specific nouns in the articles. 

```{r}
hist(lengths(V(g)$features))
```

The attribute target is average monthly traffic (October 2017 - November 2018).

```{r}
hist(log(V(g)$target, base=10))
```

Plotting the log transformation of the monthly traffic and length of word list shows no correlation

```{r}
plot(log(V(g)$target), log(lengths(V(g)$features)))
```




## Creating Subgraphs

### First Attempt at creating a subgraph

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30426)
subgraph_attempt <- induced_subgraph(g, sample(V(g), 200))

plot(subgraph_attempt,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)

sum(degree(subgraph_attempt)==0)
sum(degree(g)==0)
```


### Subgraph 1

[back to report](#generate-induced-subgraphs)


```{r}
set.seed(30326)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg1 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```


### Plot 1

```{r}
set.seed(31126)
plot(subg1,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

### Subgraph 2

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30426)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg2 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

### Plot 2

```{r}
plot(subg2,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


### Subgraph 3

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30526)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg3 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

### Plot 3

```{r}

plot(subg3,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


### Subgraph 4

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30626)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg4 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

### Plot 4

```{r}
plot(subg4,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

## Three Different Visualizations

[back to report](#visualize-the-networks)

```{r Visualize, cache = TRUE}

library(igraph)

set.seed(444)

graphs <- list(
  subg1 = subg1,
  
  subg2 = subg2,
  
  subg3 = subg3,
  subg4 = subg4
)

for (i in names(graphs)) {
  
  # 
  simp_graph <- simplify(graphs[[i]])
  #community detection for colors
  cl <- cluster_louvain(as.undirected(simp_graph))
  V(simp_graph)$color <- membership(cl)
  
  
  #adjust size to match degree level
  V(simp_graph)$size  <- 3 + 8 * (degree(simp_graph) / max(degree(simp_graph)))
  
  #remove labels
  V(simp_graph)$label <- NA
  
  #make lines between nodes cleaner
  E(simp_graph)$color <- adjustcolor("gray50", alpha.f = 0.2)


  plot(simp_graph, layout = layout_with_fr, main = paste(i, "FR"))
  
  plot(simp_graph, layout = layout_with_kk,  main = paste(i, "KK"))
  
  
  plot(simp_graph, layout = layout_nicely,  main = paste(i, "Nicely"))


}


```


### Other network Visualizations

Here are some other visualizations of the subgraphs displaying certain attributes

```{r}
set.seed(31126)
plot(
  subg1,
  layout = layout_nicely,
  vertex.size=2*log(V(subg1)$target, base=10),
  vertex.label.cex=.4,
  edge.width=2,
  main="subg1 where size is log internet traffic"
)
set.seed(31126)
plot(
  subg1,
  layout = layout_nicely,
  vertex.size=2*log(lengths(V(subg1)$features)),
  vertex.label.cex=.4,
  edge.width=2,
  main="subg1 where size is log length of the word list"
)

```



## Community Detection

[back to report](#community-detection)

```{r}
subg <- list(subg1 = subg1, 
             subg2 = subg2,
             subg3 = subg3,
             subg4 = subg4)

und_subg <- lapply(subg, as.undirected)
```

### We choose Fast Greedy, Louvain, and Walktrap (top three methods on our lecture slide and covered in ica) to perform our community detection.

[back to report](#run-community-detection-algorithms)

```{r}
und_subg <- lapply(subg, function(x) simplify(as_undirected(x)))
```

```{r}
com_fg <- lapply(und_subg, cluster_fast_greedy)
com_l <- lapply(und_subg, cluster_louvain)
com_w <- lapply(und_subg, cluster_walktrap)
```

### And then we compare modularity

[back to report](#analyze-and-interpret-the-resulting-communities)

```{r}
mod_fg <- sapply(com_fg, modularity)
mod_l <- sapply(com_l, modularity)
mod_w <- sapply(com_w, modularity)

data.frame(
subgraph = names(subg),
fast_greedy = mod_fg, 
louvain = mod_l,
walktrap = mod_w
)

ncom_fg <- sapply(com_fg, length)
ncom_l <- sapply(com_l, length)
ncom_w <- sapply(com_w, length)

data.frame(
subgraph = names(subg),
fast_greedy = ncom_fg, 
louvain = ncom_l,
walktrap = ncom_w
)
```

### And as it was recommended in our lecture to report community size, we do so as well here.

[back to report](#analyze-and-interpret-the-resulting-communities)

```{r}
sizes(com_l[[1]])
sizes(com_l[[2]])
sizes(com_l[[3]])
sizes(com_l[[4]])
```

### And we visualize with community coloring, short eg

[back to report](#analyze-and-interpret-the-resulting-communities)

```{r}
plot(com_l[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Louvain community in subgraph 1")
plot(com_l[[2]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Louvain community in subgraph 2")
plot(com_l[[3]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Louvain community in subgraph 3")

plot(com_fg[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Fast Greedy community in subgraph 1")
plot(com_fg[[1]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Fast Greedy community in subgraph 2")
plot(com_fg[[1]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Fast Greedy community in subgraph 3")

plot(com_w[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Walktrap community in subgraph 1")
plot(com_w[[2]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Walktrap community in subgraph 2")
plot(com_w[[3]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Walktrap community in subgraph 3")
```


## Adjacency Matrix

[back to report](#adjacency-matrix)

### Unordered Adjacency Matrix

```{r}
library(Matrix)

non_ord <- com_l[[1]]$membership
A <- as_adj(und_subg[[1]])
image(Matrix(A))
```

### Ordered Adjacency Matrix

```{r}
#library(Matrix)

ord <- order(com_l[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]))

ord <- order(com_fg[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]), main="subg1 adj matrix, fast greedy communities")

ord <- order(com_w[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]), main="subg1 adj matrix, walktrap communities")
```

## Analyzing Network Metrics (Degree)

[back to report](#degree)

```{r}
hist(degree(g))
```
```{r}
# Degree of Nodes in Subgraph 1
degree <- degree(subg1, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg1)$color <- "maroon"
V(subg1)$frame.color <- "white"


# Highlights top 5 node in terms of degree
V(subg1)$color[V(subg1)[head(degree_sort$ix)]] <- "#F0DC16"


plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```
```{r}
# Degree of Nodes in Subgraph 2
degree <- degree(subg2, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg2)$color <- "maroon"
V(subg2)$frame.color <- "white"

# Highlights top 5 node in terms of degree
V(subg2)$color[V(subg2)[head(degree_sort$ix)]] <- "#F0DC16"


plot(subg2,
     layout = layout_with_kk(subg1),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 3
degree <- degree(subg3, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg3)$color <- "maroon"
V(subg3)$frame.color <- "white"


# Highlights top 5 node in terms of degree
V(subg3)$color[V(subg3)[head(degree_sort$ix)]] <- "#F0DC16"

plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 4
degree <- degree(subg4, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg4)$color <- "maroon"
V(subg4)$frame.color <- "white"


# Highlights top 5 node in terms of degree
V(subg4)$color[V(subg4)[head(degree_sort$ix)]] <- "#F0DC16"

plot(subg4,
     layout = layout_with_kk(subg4),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```


## Analyzing Network Metrics (EigenCentrality)

[back to report](#eigen-metric)

```{r EigenCentrality, cache=TRUE}
print("Eigen Centrality SubGraph 1", quote=FALSE)
eig_subg1 <- eigen_centrality(subg1)$vector
hist(eig_subg1)
hist(15*sqrt(sqrt(eig_subg1)))
plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = 15*sqrt(sqrt(eig_subg1)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 2", quote=FALSE)
eig_subg2 <- eigen_centrality(subg2)$vector
hist(eig_subg2)
plot(subg2,
     layout = layout_with_kk(subg2),
     vertex.size = 15*sqrt(sqrt(eig_subg2)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 3", quote=FALSE)
eig_subg3 <- eigen_centrality(subg3)$vector
hist(eig_subg3)
plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = 15*sqrt(sqrt(eig_subg3)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 4", quote=FALSE)
eig_subg4 <- eigen_centrality(subg4)$vector
hist(eig_subg4)
plot(subg4,
     layout = layout_with_kk(subg4),
     vertex.size = 15*sqrt(sqrt(eig_subg4)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

## Monte Carlo Modularity

[back to report](#monte-carlo-estimate-of-modularity)

```{r, eval=FALSE}
set.seed(123) # For reproducibility
n_sim <- 1000
null_mods <- list(numeric(n_sim),numeric(n_sim),numeric(n_sim),numeric(n_sim))

for(i in 1:n_sim) {
  # Create a randomized graph preserving degree distribution
  # "degseq" method keeps the degree of each node the same
  rg1 <- sample_degseq(degree(subg1), method = "vl")
  rg2 <- sample_degseq(degree(subg2), method = "vl")
  rg3 <- sample_degseq(degree(subg3), method = "vl")
  rg4 <- sample_degseq(degree(subg4), method = "vl")
  
  # Run the SAME community detection algorithm on the random graph
  r_comm1 <- cluster_louvain(rg1)
  r_comm2 <- cluster_louvain(rg2)
  r_comm3 <- cluster_louvain(rg3)
  r_comm4 <- cluster_louvain(rg4)
  
  null_mods[[1]][i] <- modularity(r_comm1)
  null_mods[[2]][i] <- modularity(r_comm2)
  null_mods[[3]][i] <- modularity(r_comm3)
  null_mods[[4]][i] <- modularity(r_comm4)
}

save(null_mods, file="null_mods.rda")
```

```{r}
load('null_mods.rda')

hist(null_mods[[1]], main="Modularity of random graph with connectivity of subg1",
     xlim=c(min(null_mods[[1]]), max(0.2477544, max(null_mods[[1]]))))

abline(v=0.2477544, col="red", lwd=3, lty=2)
legend("topright", legend=c("Observed", "Null Dist"), 
       col=c("red", "black"), lty=c(2,1), lwd=c(3,1))

hist(null_mods[[2]], main="Modularity of random graph with connectivity of subg2",
     xlim=c(min(null_mods[[2]]), max(0.3608913, max(null_mods[[2]]))))

abline(v=0.3608913, col="red", lwd=3, lty=2)
legend("topright", legend=c("Observed", "Null Dist"), 
       col=c("red", "black"), lty=c(2,1), lwd=c(3,1))

hist(null_mods[[3]], main="Modularity of random graph with connectivity of subg3",
     xlim=c(min(null_mods[[3]]), max(0.2025717, max(null_mods[[3]]))))

abline(v=0.2025717, col="red", lwd=3, lty=2)
legend("topright", legend=c("Observed", "Null Dist"), 
       col=c("red", "black"), lty=c(2,1), lwd=c(3,1))

hist(null_mods[[4]], main="Modularity of random graph with connectivity of subg4",
     xlim=c(min(null_mods[[4]]), max(0.3141144, max(null_mods[[4]]))))

abline(v=0.3141144, col="red", lwd=3, lty=2)
legend("topright", legend=c("Observed", "Null Dist"), 
       col=c("red", "black"), lty=c(2,1), lwd=c(3,1))
```

## Analyzing Network Metrics (Betweenness)

[back to report](#betweenness)

```{r Betweenness, cache=TRUE}
print("Betweenness SubGraph 1", quote=FALSE)
bet_subg1 <- betweenness(subg1)
hist(bet_subg1)
plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = 2*log(bet_subg1+1),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Betweenness SubGraph 2", quote=FALSE)
bet_subg2 <- betweenness(subg2)
hist(bet_subg2)
plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = 2*log(bet_subg1+1),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Betweenness SubGraph 3", quote=FALSE)
bet_subg3 <- betweenness(subg3)
hist(bet_subg3)
plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = 2*log(bet_subg3+1),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Betweenness SubGraph 4", quote=FALSE)
bet_subg4 <- betweenness(subg4)
hist(bet_subg4)
plot(subg1,
     layout = layout_with_kk(subg4),
     vertex.size = 2*log(bet_subg4+1),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```


## Analyzing Network Metrics (Closeness)

[back to report](#closeness)

```{r Closeness, cache=TRUE}
print("Closeness SubGraph 1", quote=FALSE)
close_subg1 <- closeness(subg1)
hist(close_subg1)
plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = 4000*close_subg1,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Closeness SubGraph 2", quote=FALSE)
close_subg2 <- closeness(subg2)
hist(close_subg2)
plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = 4000*close_subg2,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Closeness SubGraph 3", quote=FALSE)
close_subg3 <- closeness(subg3)
hist(close_subg3)
plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = 4000*close_subg3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Closeness SubGraph 4", quote=FALSE)
close_subg4 <- closeness(subg4)
hist(close_subg4)
plot(subg1,
     layout = layout_with_kk(subg4),
     vertex.size = 4000*close_subg4,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```
