---
title: "PSTAT 194 Individual Final Messing Around"
author: "Ayden Kelleher"
date: "2026-03-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
    highlight: tango
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load and Prepare the Data

## Load your chosen dataset into the course RStudio server and convert it to an ‘igraph’ object.

We loaded the data in [Code Snippet 1](#snippet-1)


## Consider what the vertices and edges represent in your network.

In our network, the vertices represent wikipedia articles, and edges represent mutual links between them. Since we are using the musae squirrel network, our articles are about topics related to squirrels, and the edges are mutual links between them. 


## Identify basic network elements: vertices, edges, weights, directionality, etc

Using code [snippet 2](#snippet-2), we can see that we have 5201 nodes, 217073 edges, that our graph is unweighted, and that our graph is undirected. 

# Generate Induced Subgraphs:

## Create four induced subgraphs with 200 random nodes each from your dataset.

To create our induced subgraphs, we first attempted to [sample 200](#first-attempt-at-creating-a-subgraph) with 200 random nodes from our graph. However, we ran into the issue that much of our graph had degree zero, with 79 out of 200 nodes being unconnected to any other node, in contrast to our entire graph g have no points that are of degree zero. Thus any random sample would lose some structure. 

To get around this, what we did instead was to draw a sample of 40 different nodes. Then we took a sample of 5 different neighbors of the nodes, or all of the neighbors if they had less than 5 total neighbors. Then we take 200 of those nodes, and put it in a subgraph. This subgraph is a biased subgraph, but it preserves some of the structure of the graph. We did this in each of the subgraphs, which can be seen in [subgraph 1](subgraph-1), [subgraph 2](subgraph-2), [subgraph 3](subgraph-3), and [subgraph 4](subgraph-4)

Using the basic plotting, we plotted using the default layouts.

# Visualize the Networks:

## Use three different layouts from the igraph package to visualize each network.

## Experiment with node and edge attributes to enhance visualizations.

# Analyze Network Metrics:

## Explore various network, node, and edge metrics. Interpret these metrics in the context of your network.

### Degree

### Closeness

### Betweenness

### Eigen Metric

# Community Detection:

## Run community detection algorithms.

## Analyze and interpret the resulting communities

# Adjacency Matrix:

## Examine the adjacency matrix of your network.

## Reorder vertices to highlight patterns.

# Reorder Nodes:

## Reorder nodes based on community or connected component.

## Observe changes in the adjacency matrix and identify clearer patterns

# Summarize Findings:

## Prepare a presentation and a short write-up summarizing your analysis and findings.

# Appendix

## Snippet 1

```{r}
#install.packages('igraph')
library(igraph)
#install.packages('tidygraph')
library(tidygraph)

edges_matrix <- as.matrix(read.csv("musae_squirrel_edges.csv"))

edges_matrix[,1] <- factor(edges_matrix[,1]) # if we dont change this to factor, then there are too many id names idk. If you don't believe me comment this out :/

g <- graph_from_edgelist(edges_matrix, directed = FALSE)

```

```{r}
length(V(g)) #this is exactly what the help page says
length(E(g)) #this is quite a few more edges than the help page says
```

## Snippet 2

```{r}
length(V(g))
length(E(g))
is_weighted(g)
is_directed(g)
summary(g)
```


## First Attempt at creating a subgraph

```{r}
set.seed(30426)
subgraph_attempt <- induced_subgraph(g, sample(V(g), 200))

plot(subgraph_attempt,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)

sum(degree(subgraph_attempt)==0)
sum(degree(g)==0)
```


## Subgraph 1




```{r}
set.seed(30326)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg1 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```


## Plot 1

```{r}
plot(subg1,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

## Subgraph 2

```{r}
set.seed(30426)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg2 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

## Plot 2

```{r}
plot(subg2,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


## Subgraph 3

```{r}
set.seed(30526)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg3 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

## Plot 3

```{r}

plot(subg3,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


## Subgraph 4

```{r}
set.seed(30626)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg4 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

## Plot 4

```{r}
plot(subg4,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

# Community Detection
```{r}

library(igraph)

set.seed(444)

graphs <- list(
  subg1 = subg1,
  
  subg2 = subg2,
  
  subg3 = subg3,
  subg4 = subg4
)

for (i in names(graphs)) {
  
  # 
  simp_graph <- simplify(graphs[[i]])
  #community detection for colors
  cl <- cluster_louvain(as.undirected(simp_graph))
  V(simp_graph)$color <- membership(cl)
  
  
  #adjust size to match degree level
  V(simp_graph)$size  <- 3 + 8 * (degree(simp_graph) / max(degree(simp_graph)))
  
  #remove labels
  V(simp_graph)$label <- NA
  
  #make lines between nodes cleaner
  E(simp_graph)$color <- adjustcolor("gray50", alpha.f = 0.2)


  plot(simp_graph, layout = layout_with_fr, main = paste(i, "FR"))
  
  plot(simp_graph, layout = layout_with_kk,  main = paste(i, "KK"))
  
  
  plot(simp_graph, layout = layout_nicely,  main = paste(i, "Nicely"))


}


```

```{r}
subg <- list(subg1 = subg1, 
             subg2 = subg2,
             subg3 = subg3,
             subg4 = subg4)

und_subg <- lapply(subg, as.undirected)
```

## We choose Fast Greedy, Louvain, and Walktrap (top three methods on our lecture slide and covered in ica) to perform our community detection.

```{r}
und_subg <- lapply(subg, function(x) simplify(as_undirected(x)))
```

```{r}
com_fg <- lapply(und_subg, cluster_fast_greedy)
com_l <- lapply(und_subg, cluster_louvain)
com_w <- lapply(und_subg, cluster_walktrap)
```

## And then we compare modularity

```{r}
mod_fg <- sapply(com_fg, modularity)
mod_l <- sapply(com_l, modularity)
mod_w <- sapply(com_w, modularity)

data.frame(
subgraph = names(subg),
fast_greedy = mod_fg, 
louvain = mod_l,
walktrap = mod_w
)

ncom_fg <- sapply(com_fg, length)
ncom_l <- sapply(com_l, length)
ncom_w <- sapply(com_w, length)

data.frame(
subgraph = names(subg),
fast_greedy = ncom_fg, 
louvain = ncom_l,
walktrap = ncom_w
)
```
## And as it was recommended in our lecture to report community size, we do so as well here.
```{r}
sizes(com_l[[1]])
sizes(com_l[[2]])
sizes(com_l[[3]])
sizes(com_l[[4]])
```
## And we visualize with community coloring, short eg

```{r}
plot(com_l[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Louvain community in subgraph 1")
plot(com_l[[2]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Louvain community in subgraph 2")
plot(com_l[[3]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Louvain community in subgraph 3")
```


# Adjacency Matrix
```{r}
library(Matrix)

ord <- order(com_l[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]))
```

## Analyzing Network Metrics (Degree)
```{r}
# Degree of Nodes in Subgraph 1
degree <- degree(subg1, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg1)$color <- "white"
V(subg1)$label.color <- "black"
V(subg1)$frame.color <- "black"

V(subg1)$color[V(subg1)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```
```{r}
# Degree of Nodes in Subgraph 2
degree <- degree(subg2, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg2)$color <- "white"
V(subg2)$label.color <- "black"
V(subg2)$frame.color <- "black"

V(subg2)$color[V(subg2)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg2,
     layout = layout_with_kk(subg2),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 3
degree <- degree(subg3, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg3)$color <- "white"
V(subg3)$label.color <- "black"
V(subg3)$frame.color <- "black"

V(subg3)$color[V(subg3)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 4
degree <- degree(subg4, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg4)$color <- "white"
V(subg4)$label.color <- "black"
V(subg4)$frame.color <- "black"

V(subg4)$color[V(subg4)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg4,
     layout = layout_with_kk(subg4),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```


