---
title: "PSTAT 194 Individual Final Messing Around"
author: "Ayden Kelleher"
date: "2026-03-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
    highlight: tango
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load and Prepare the Data

## Load your chosen dataset into the course RStudio server and convert it to an ‘igraph’ object.

We loaded the data in [Code Snippet 1](#snippet-1)


## Consider what the vertices and edges represent in your network.

In our network, the vertices represent wikipedia articles, and edges represent mutual links between them. Since we are using the musae squirrel network, our articles are about topics related to squirrels, and the edges are mutual links between them. 


## Identify basic network elements: vertices, edges, weights, directionality, etc

Using code [snippet 2](#snippet-2), we can see that we have 5201 nodes, 217073 edges, that our graph is unweighted, and that our graph is undirected. 

# Generate Induced Subgraphs:

## Create four induced subgraphs with 200 random nodes each from your dataset.

To create our induced subgraphs, we first attempted to [sample 200](#first-attempt-at-creating-a-subgraph) with 200 random nodes from our graph. However, we ran into the issue that much of our graph had degree zero, with 79 out of 200 nodes being unconnected to any other node, in contrast to our entire graph g have no points that are of degree zero. Thus any random sample would lose some structure. 

To get around this, what we did instead was to draw a sample of 40 different nodes. Then we took a sample of 5 different neighbors of the nodes, or all of the neighbors if they had less than 5 total neighbors. Then we take 200 of those nodes, and put it in a subgraph. This subgraph is a biased subgraph, but it preserves some of the structure of the graph. We did this in each of the subgraphs, which can be seen in [subgraph 1](subgraph-1), [subgraph 2](subgraph-2), [subgraph 3](subgraph-3), and [subgraph 4](subgraph-4)

Using the basic plotting, we plotted using the default layouts.

# Visualize the Networks:

## Use three different layouts from the igraph package to visualize each network.

For our visualizations, we chose 3 different layouts, the Fruchterman-Reingold layout, the Kamada-Kawai layout, and the layout nicely, where the plots can be found [here](#three-different-visualizations). 

## Experiment with node and edge attributes to enhance visualizations.

To enhance out visualizations, we chose to do a few different things

  - adjust the node size by the following function of degree `3 + 8 * (degree(simp_graph) / max(degree(simp_graph)))`
  - color in nodes based on louvain communities to make the groupings more obvious
  - remove the node labels. Our nodes are labelled 0 to 5200. This gives us no information about the corresponding articles, so not showing the degree label doesn't lose any information. Even if we had the article names, it would be very difficult to read.
  - Make the lines between the nodes more see through so they take up less of the visual space.

# Analyze Network Metrics:

## Explore various network, node, and edge metrics. Interpret these metrics in the context of your network.

### Degree

The degree of a node is the measure of amount of edges connecting to a node. For directed graphs, we can have in and out degree, but our graph is undirected, so we don't need to worry about that. The first thing we will do is to look at the [histogram of degree](#analyzing-network-metrics-degree) of the entire network. Here we can see that the distribution is heavily skewed to the right, with a giant peak at $[0,100]$. 

When we look at the individual subgraphs, we can see a similar pattern. Each subgraph is skewed to the right. The most connected nodes have a bit less than 100 nodes, which is a little less than half the network. This tracks with the most connected nodes in the full network being connected to about half of the other nodes. One interesting pattern that was not expected is that for each of the subgraphs, there is a smaller peak at $[30,50]$. That is to say for each of the subgraphs degree distribution, there was a peak at zero, it decreased reaching a low at around $20$, then hit a second peak somewhere in $[30,50]$. We wonder if this is a result of a true pattern in the data, or if it is an artifact from the way we sampled. Recall we randomly sampled 40 nodes, then sampled neighbors of those nodes. Since nodes of a higher degree are connected to more nodes, they are more likely to appear when we sample neighbors of the nodes, so it is reasonable to infer this is a result of the way we sampled our data. 

### Closeness

### Betweenness

### Eigen Metric

The eigen centrality metric is a measure of centrality that, according to the documentation, states "eigenvector centrality scores correspond to the values of the principal eigenvector of the graph's adjacency matrix; these scores may, in turn, be interpreted as arising from a reciprocal process in which the centrality of each actor is proportional to the sum of the centralities of those actors to whom he or she is connected. In general, vertices with high eigenvector centralities are those which are connected to many other vertices which are, in turn, connected to many others (and so on)" from [here](https://r.igraph.org/reference/eigen_centrality.html). The thing to keep in mind is that the score gets bigger when a node is connected to more nodes, and these nodes are connected to more nodes, and so on. 

Looking at the plots of subgraphs [here](#analyzing-network-metrics-eigencentrality), we will start with the histograms. We can see that eigen centrality is distributed on $[0,1]$, where 1 is more connected. Looking at our subgraphs, we see the mode is at $[0,.1]$, with a second, much smaller, peak at about $[.4,.5]$. Comparing these to our degree distributions, we can see their shape is similar, with a main peak at 0, and a second smaller peak towards the middle of the data, which makes sense because both these node metrics are correlated. To show how this data looks when plotting the whole network, we cannot use the raw eigen metric because our nodes are too small, so we will use a transformation of the data. To give the reader an idea what this transformation does, we also plotted the histogram of the transformed data for the first subgraph.

Looking at the actual plots, we can see there there are a bunch of central nodes that are grouped together, and then there are some less connected peripheral nodes, that are very small. This is most prominant on subgraph 3. This shows eigen centrality is a good measure of centrality.

# Community Detection:

## Run community detection algorithms.

## Analyze and interpret the resulting communities

# Adjacency Matrix:

## Examine the adjacency matrix of your network.

## Reorder vertices to highlight patterns.

# Reorder Nodes:

## Reorder nodes based on community or connected component.

## Observe changes in the adjacency matrix and identify clearer patterns

# Summarize Findings:

## Prepare a presentation and a short write-up summarizing your analysis and findings.

# Appendix

## Snippet 1

```{r}
#install.packages('igraph')
library(igraph)
#install.packages('tidygraph')
library(tidygraph)

edges_matrix <- as.matrix(read.csv("musae_squirrel_edges.csv"))

edges_matrix[,1] <- factor(edges_matrix[,1]) # if we dont change this to factor, then there are too many id names idk. If you don't believe me comment this out :/

g <- graph_from_edgelist(edges_matrix, directed = FALSE)

```

```{r}
length(V(g)) #this is exactly what the help page says
length(E(g)) #this is quite a few more edges than the help page says
```

## Snippet 2

```{r}
length(V(g))
length(E(g))
is_weighted(g)
is_directed(g)
summary(g)
```


## First Attempt at creating a subgraph

```{r}
set.seed(30426)
subgraph_attempt <- induced_subgraph(g, sample(V(g), 200))

plot(subgraph_attempt,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)

sum(degree(subgraph_attempt)==0)
sum(degree(g)==0)
```


## Subgraph 1




```{r}
set.seed(30326)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg1 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```


## Plot 1

```{r}
plot(subg1,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

## Subgraph 2

```{r}
set.seed(30426)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg2 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

## Plot 2

```{r}
plot(subg2,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


## Subgraph 3

```{r}
set.seed(30526)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg3 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

## Plot 3

```{r}

plot(subg3,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


## Subgraph 4

```{r}
set.seed(30626)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg4 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

## Plot 4

```{r}
plot(subg4,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

# Three Different Visualizations

```{r Visualize, cache = TRUE}

library(igraph)

set.seed(444)

graphs <- list(
  subg1 = subg1,
  
  subg2 = subg2,
  
  subg3 = subg3,
  subg4 = subg4
)

for (i in names(graphs)) {
  
  # 
  simp_graph <- simplify(graphs[[i]])
  #community detection for colors
  cl <- cluster_louvain(as.undirected(simp_graph))
  V(simp_graph)$color <- membership(cl)
  
  
  #adjust size to match degree level
  V(simp_graph)$size  <- 3 + 8 * (degree(simp_graph) / max(degree(simp_graph)))
  
  #remove labels
  V(simp_graph)$label <- NA
  
  #make lines between nodes cleaner
  E(simp_graph)$color <- adjustcolor("gray50", alpha.f = 0.2)


  plot(simp_graph, layout = layout_with_fr, main = paste(i, "FR"))
  
  plot(simp_graph, layout = layout_with_kk,  main = paste(i, "KK"))
  
  
  plot(simp_graph, layout = layout_nicely,  main = paste(i, "Nicely"))


}


```


# Community Detection

```{r}
subg <- list(subg1 = subg1, 
             subg2 = subg2,
             subg3 = subg3,
             subg4 = subg4)

und_subg <- lapply(subg, as.undirected)
```

## We choose Fast Greedy, Louvain, and Walktrap (top three methods on our lecture slide and covered in ica) to perform our community detection.

```{r}
und_subg <- lapply(subg, function(x) simplify(as_undirected(x)))
```

```{r}
com_fg <- lapply(und_subg, cluster_fast_greedy)
com_l <- lapply(und_subg, cluster_louvain)
com_w <- lapply(und_subg, cluster_walktrap)
```

## And then we compare modularity

```{r}
mod_fg <- sapply(com_fg, modularity)
mod_l <- sapply(com_l, modularity)
mod_w <- sapply(com_w, modularity)

data.frame(
subgraph = names(subg),
fast_greedy = mod_fg, 
louvain = mod_l,
walktrap = mod_w
)

ncom_fg <- sapply(com_fg, length)
ncom_l <- sapply(com_l, length)
ncom_w <- sapply(com_w, length)

data.frame(
subgraph = names(subg),
fast_greedy = ncom_fg, 
louvain = ncom_l,
walktrap = ncom_w
)
```

## And as it was recommended in our lecture to report community size, we do so as well here.

```{r}
sizes(com_l[[1]])
sizes(com_l[[2]])
sizes(com_l[[3]])
sizes(com_l[[4]])
```

## And we visualize with community coloring, short eg

```{r}
plot(com_l[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Louvain community in subgraph 1")
plot(com_l[[2]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Louvain community in subgraph 2")
plot(com_l[[3]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Louvain community in subgraph 3")

plot(com_fg[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Fast Greedy community in subgraph 1")
plot(com_fg[[1]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Fast Greedy community in subgraph 2")
plot(com_fg[[1]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Fast Greedy community in subgraph 3")

plot(com_w[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Walktrap community in subgraph 1")
plot(com_w[[2]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Walktrap community in subgraph 2")
plot(com_w[[3]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Walktrap community in subgraph 3")
```


# Adjacency Matrix

```{r}
library(Matrix)

ord <- order(com_l[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]))
```

## Analyzing Network Metrics (Degree)

[back to report](#degree)

```{r}
hist(degree(g))
```

```{r}
# Degree of Nodes in Subgraph 1
degree <- degree(subg1, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg1)$color <- "white"
V(subg1)$label.color <- "black"
V(subg1)$frame.color <- "black"

V(subg1)$color[V(subg1)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```
```{r}
# Degree of Nodes in Subgraph 2
degree <- degree(subg2, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg2)$color <- "white"
V(subg2)$label.color <- "black"
V(subg2)$frame.color <- "black"

V(subg2)$color[V(subg2)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg2,
     layout = layout_with_kk(subg2),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 3
degree <- degree(subg3, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg3)$color <- "white"
V(subg3)$label.color <- "black"
V(subg3)$frame.color <- "black"

V(subg3)$color[V(subg3)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 4
degree <- degree(subg4, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
sort(degree, decreasing = T, index.return = T)
V(subg4)$color <- "white"
V(subg4)$label.color <- "black"
V(subg4)$frame.color <- "black"

V(subg4)$color[V(subg4)[c(29,37,77,62,52)]] <- "#F0DC16"

plot(subg4,
     layout = layout_with_kk(subg4),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

## Analyzing Network Metrics (EigenCentrality)

[back to report](#eigen-metric)

```{r EigenCentrality, cache=TRUE}
print("Eigen Centrality SubGraph 1", quote=FALSE)
eig_subg1 <- eigen_centrality(subg1)$vector
hist(eig_subg1)
hist(15*sqrt(sqrt(eig_subg1)))
plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = 15*sqrt(sqrt(eig_subg1)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 2", quote=FALSE)
eig_subg2 <- eigen_centrality(subg2)$vector
hist(eig_subg2)
plot(subg2,
     layout = layout_with_kk(subg2),
     vertex.size = 15*sqrt(sqrt(eig_subg2)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 3", quote=FALSE)
eig_subg3 <- eigen_centrality(subg3)$vector
hist(eig_subg3)
plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = 15*sqrt(sqrt(eig_subg3)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 4", quote=FALSE)
eig_subg4 <- eigen_centrality(subg4)$vector
hist(eig_subg4)
plot(subg4,
     layout = layout_with_kk(subg4),
     vertex.size = 15*sqrt(sqrt(eig_subg4)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```