---
title: "PSTAT 194 Final Project"
author: "Ayden Kelleher"
date: "2026-03-04"
output:
  word_document:
    toc: true
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction:

For our network analysis, we chose for our network to be the Multi-scale Attributed Node Embedding Network Wikipedia page-page networks on Squirrels. This network consists of a subset of wikipedia articles, where the nodes are wikipedia articles on squirrels. The edges are mutual hyperlinks between articles, thus our graph is undirected. Using the `summary` function, we can see that our network has 5201 nodes, 217073 edges, that our graph is unweighted, and that our graph is undirected. Our graph has 2 node attribute. The first one, called target, is the average monthly traffic (October 2017 - November 2018). The second one, called features in our data, is a list containing a collection of numbers. These numbers correspond to specific nouns in the articles, but we couldn't find which nouns go to what number. We have plotted histograms of the attributes [here](#snippet-2) to give a basic idea of what the distribution of these attributes are. 

The objective our analysis is to see what the structure of wikipedia articles looks like, and to see if we can get an idea of what attribute influence how much traffic a node gets. 

# Methodology:

## Steps followed in data processing.

The first thing we did was we loaded a few `r` packages, including `igraph`, `tidygraph`, `jsonlite`, and `tidyr`, along with the data in [Code Snippet 1](#snippet-1). The graph data we had was an edge list, which we turned into a graph by turning it into a matrix, then turning this matrix into a graph using the `graph_from_edgelist` from the `igraph` package. We loaded the json data using the `fromJSON` function from the `jsonlite` package. Lastly, we loaded in the target data using the `read.csv` function from base `r`. We then added these as vertex attributes. 

# Generate Induced Subgraphs:

## Create four induced subgraphs with 200 random nodes each from your dataset.

To create our induced subgraphs, we first attempted to [sample 200](#first-attempt-at-creating-a-subgraph) with 200 random nodes from our graph. However, we ran into the issue that much of our graph had degree zero, with 79 out of 200 nodes being unconnected to any other node, in contrast to our entire graph g have no points that are of degree zero. Thus any random sample would lose some structure. 

To get around this, what we did instead was to draw a sample of 40 different nodes. Then we took a sample of 5 different neighbors of the nodes, or all of the neighbors if they had less than 5 total neighbors. Then we take 200 of those nodes, and put it in a subgraph. This subgraph is a biased subgraph, but it preserves some of the structure of the graph. We did this in each of the subgraphs, which can be seen in [subgraph 1](#subgraph-1), [subgraph 2](#subgraph-2), [subgraph 3](#subgraph-3), and [subgraph 4](#subgraph-4)

Using the basic plotting, we plotted using the default layouts.

# Visualize the Networks:

## Use three different layouts from the igraph package to visualize each network.

For our visualizations, we chose 3 different layouts, the Fruchterman-Reingold layout, the Kamada-Kawai layout, and the layout nicely, where the plots can be found [here](#three-different-visualizations). 

## Experiment with node and edge attributes to enhance visualizations.

To enhance out visualizations, we chose to do a few different things

  - adjust the node size by the following function of degree `3 + 8 * (degree(simp_graph) / max(degree(simp_graph)))`
  - color in nodes based on louvain communities to make the groupings more obvious
  - remove the node labels. Our nodes are labelled 0 to 5200. This gives us no information about the corresponding articles, so not showing the degree label doesn't lose any information. Even if we had the article names, it would be very difficult to read.
  - Make the lines between the nodes more see through so they take up less of the visual space.

# Analyze Network Metrics:

## Explore various network, node, and edge metrics. Interpret these metrics in the context of your network.

### Degree

The degree of a node is the measure of amount of edges connecting to a node. For directed graphs, we can have in and out degree, but our graph is undirected, so we don't need to worry about that. The first thing we will do is to look at the [histogram of degree](#analyzing-network-metrics-degree) of the entire network. Here we can see that the distribution is heavily skewed to the right, with a giant peak at $[0,100]$. 

When we look at the individual subgraphs, we can see a similar pattern. Each subgraph is skewed to the right. The most connected nodes have a bit less than 100 nodes, which is a little less than half the network. This tracks with the most connected nodes in the full network being connected to about half of the other nodes. One interesting pattern that was not expected is that for each of the subgraphs, there is a smaller peak at $[30,50]$. That is to say for each of the subgraphs degree distribution, there was a peak at zero, it decreased reaching a low at around $20$, then hit a second peak somewhere in $[30,50]$. We wonder if this is a result of a true pattern in the data, or if it is an artifact from the way we sampled. Recall we randomly sampled 40 nodes, then sampled neighbors of those nodes. Since nodes of a higher degree are connected to more nodes, they are more likely to appear when we sample neighbors of the nodes, so it is reasonable to infer this is a result of the way we sampled our data. 

### Closeness

### Betweenness

### Eigen Metric

The eigen centrality metric is a measure of centrality that, according to the documentation, states "eigenvector centrality scores correspond to the values of the principal eigenvector of the graph's adjacency matrix; these scores may, in turn, be interpreted as arising from a reciprocal process in which the centrality of each actor is proportional to the sum of the centralities of those actors to whom he or she is connected. In general, vertices with high eigenvector centralities are those which are connected to many other vertices which are, in turn, connected to many others (and so on)" from [here](https://r.igraph.org/reference/eigen_centrality.html). The thing to keep in mind is that the score gets bigger when a node is connected to more nodes, and these nodes are connected to more nodes, and so on. 

Looking at the plots of subgraphs [here](#analyzing-network-metrics-eigencentrality), we will start with the histograms. We can see that eigen centrality is distributed on $[0,1]$, where 1 is more connected. Looking at our subgraphs, we see the mode is at $[0,.1]$, with a second, much smaller, peak at about $[.4,.5]$. Comparing these to our degree distributions, we can see their shape is similar, with a main peak at 0, and a second smaller peak towards the middle of the data, which makes sense because both these node metrics are correlated. To show how this data looks when plotting the whole network, we cannot use the raw eigen metric because our nodes are too small, so we will use a transformation of the data. To give the reader an idea what this transformation does, we also plotted the histogram of the transformed data for the first subgraph.

Looking at the actual plots, we can see there there are a bunch of central nodes that are grouped together, and then there are some less connected peripheral nodes, that are very small. This is most prominent on subgraph 3. This shows eigen centrality is a good measure of centrality.

# Community Detection:

## Run community detection algorithms

We chose to run [three different community detection algorithms](#we-choose-fast-greedy-louvain-and-walktrap-top-three-methods-on-our-lecture-slide-and-covered-in-ica-to-perform-our-community-detection.), fast greedy, louvain, and walktrap. 

The way fast greedy works is by first declaring each node its own community, then it combines nodes greedily, that is to say maximizing the increase in modularity for each step, and stops when it can no longer increase the modularity. One thing to look out for is when working with larger networks, it can fail to find smaller communities .

The way the louvain community detection works is by declaring each node its own community, then moves nodes into the community that increases the modularity the most, and then collapses the communities into super nodes. 
The way walktrap works is by starting a random walk in to run short random walks. The idea is that nodes in the same community will appear more on these short random walks. 

## Analyze and interpret the resulting communities

Comparing the results from the algorithms was interesting. Starting with the [modularity of our subgraphs](#and-then-we-compare-modularity), none of the graphs had very large modularity. Below is the results

$$\begin{array} &\text{subgraph #}&\text{fast greedy}&\text{louvain}&\text{walktrap} \\
 1& 0.1911257 &0.2461304 &0.15813260 \\
2& 0.3522581 &0.3608913 &0.29551518 \\
3& 0.1545764 &0.2025717 &0.10072104 \\
4& 0.2772735 &0.3141144 &0.08605815
\end{array}$$

In class, we learned that modularity over $.3$ suggests meaningful community structure. However, only subgraphs 2 and 4 had any communities with modularity at least this high. However, subgraphs 1 and 3 still had modularity $.246$ and $.202$ respectively most modular community structrues, which isn't that much smaller than $.3$. This suggests that our network has a weak community structure. 

Looking at the number of individual communities for each subgraph, we can see the following. 

$$\begin{array}
& \text{subgraph #}& \text{fast greedy}& \text{louvain}& \text{walktrap} \\
1 &6       &5       &31 \\
2 &6       &6       &27 \\
3 &4       &7       &27 \\
4 &6       &5       &20
\end{array}$$

This gives us the biggest differences between the different community finding algorithms. We can see that both the fast greedy and the louvain algorithms ended up with a small amount of communities, around 6 for each subgraph. For the walktrap, each subgraph had many more communities, with 31, 27, 27, and 20 for each of the subgraphs respectively. Looking at [the plots](#and-we-visualize-with-community-coloring-short-eg) of the subgraphs with the community coloring, we can see that the walktrap has many isolated nodes that count as their own communities. This tells us if a node is relatively disconnected from the rest of the network, the walktrap will have a hard time placing it in a larger community. 


# Adjacency Matrix:

## Examine the adjacency matrix of your network.

We felt it was important to show at the unordered adjacency matrix to show what an uncorrelated adjacency matrix looks like. Looking at the [unordered adjacency matrix](#unordered-adjacency-matrix) for subgraph 1 using the louvain method, we can see very little pattern. The most of a pattern one could make out is that there a few horizontal lines, representing the nodes that are most connected, and a few blocks, representing multiple nodes connected to similar nodes. This lack of pattern is expected since our nodes have not been ordered yet. 

## Reorder vertices to highlight patterns.

After [reordering our nodes](#ordered-adjacency-matrix), we can see a clearer pattern. There are 5 squares, filled in by varying amounts, on the diagnol. This matches the number of communities subgraph 1 has according to the louvain method, which confirms our observation of 5 squares. Some observations we made were that 

- The first square is much more filled in than any of the other squares. 
- We can see many horizontal/vertical lines, especially the areas overlapping with the nodes of the first community. This shows that the first community has a higher average degree than the other communities. 
- The other 4 communities have a much lower degree on average, and are much less connected among themselves, and have lots of empty space in their squares. This shows that our subgraph is quite sparse. 

Our subgraph consisted on a single connected component. Thus we cannot reorder our network by connected components. 

One thing we wonder is our sampling strategy effected this result. Recall we selected 40 random nodes, then for each node, selected either 5 neighbors, or if they didn't have 5 neighbors, all of their neighbors. Then we selected 200 of those nodes. Thus, most of the nodes we chose were not quite randomly chosen, but conditioned on being neighbors with a random node, which biases the nodes we chose to those with a higher degree. Thus our network consists of nodes that are connected to more nodes than the typical node in our network. This could lead to less communities that are larger than expected in our subgraphs. 

# Monte Carlo Estimate of Modularity

To test if our community structure was meaningful, we ran a [monte carlo](#monte-carlo-modularity) estimate of the modularity of a random network with the same amount of nodes and edges. The way we did this was by 

# Results:

## Key findings from the analysis.

## Visual and textual summaries.

# Conclusion:

## Summary of insights gained.

## Possible future directions for further analysis.

# References:

Rozemberczki, Benedek, Carl Allen, and Rik Sarkar. “Multi-Scale Attributed Node Embedding.” Journal of Complex Networks, vol. 9, no. 2, 2021. https://arxiv.org/abs/1909.13021. 


# Appendix

## Snippet 1

[back to report](#load-and-prepare-the-data)

```{r}
#install.packages('igraph')
library(igraph)
#install.packages('tidygraph')
library(tidygraph)
library(jsonlite )
library(tidyr)

edges_matrix <- as.matrix(read.csv("musae_squirrel_edges.csv"))

json_data <- fromJSON("squirrel.json")

target_data <- read.csv('squirrel_target.csv')

edges_matrix[,1] <- factor(edges_matrix[,1]) # if we dont change this to factor, then there are too many id names idk. If you don't believe me comment this out :/

g <- graph_from_edgelist(edges_matrix, directed = FALSE)

V(g)$features <- json_data

V(g)$target <- target_data$target

```

```{r}
length(V(g)) #this is exactly what the help page says
length(E(g)) #this is quite a few more edges than the help page says
```

## Snippet 2

[back to report](#identify-basic-network-elements-vertices-edges-weights-directionality-etc)

```{r}
length(V(g))
length(E(g))
is_weighted(g)
is_directed(g)
summary(g)
```

A brief look at the network features. These are specific nouns in the articles. 

```{r}
hist(lengths(V(g)$features))
```

The attribute target is average monthly traffic (October 2017 - November 2018).

```{r}
hist(log(V(g)$target, base=10))
```

Plotting the log transformation of the monthly traffic and length of word list shows no correlation

```{r}
plot(log(V(g)$target), log(lengths(V(g)$features)))
```




## Creating Subgraphs

### First Attempt at creating a subgraph

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30426)
subgraph_attempt <- induced_subgraph(g, sample(V(g), 200))

plot(subgraph_attempt,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)

sum(degree(subgraph_attempt)==0)
sum(degree(g)==0)
```


### Subgraph 1

[back to report](#generate-induced-subgraphs)


```{r}
set.seed(30326)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg1 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```


### Plot 1

```{r}
set.seed(31126)
plot(subg1,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

### Subgraph 2

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30426)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg2 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

### Plot 2

```{r}
plot(subg2,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


### Subgraph 3

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30526)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg3 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

### Plot 3

```{r}

plot(subg3,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


### Subgraph 4

[back to report](#generate-induced-subgraphs)

```{r}
set.seed(30626)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg4 <- induced_subgraph(g, unique(sampled_vertices)[1:200])
```

### Plot 4

```{r}
plot(subg4,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

## Three Different Visualizations

[back to report](#visualize-the-networks)

```{r Visualize, cache = TRUE}

library(igraph)

set.seed(444)

graphs <- list(
  subg1 = subg1,
  
  subg2 = subg2,
  
  subg3 = subg3,
  subg4 = subg4
)

for (i in names(graphs)) {
  
  # 
  simp_graph <- simplify(graphs[[i]])
  #community detection for colors
  cl <- cluster_louvain(as.undirected(simp_graph))
  V(simp_graph)$color <- membership(cl)
  
  
  #adjust size to match degree level
  V(simp_graph)$size  <- 3 + 8 * (degree(simp_graph) / max(degree(simp_graph)))
  
  #remove labels
  V(simp_graph)$label <- NA
  
  #make lines between nodes cleaner
  E(simp_graph)$color <- adjustcolor("gray50", alpha.f = 0.2)


  plot(simp_graph, layout = layout_with_fr, main = paste(i, "FR"))
  
  plot(simp_graph, layout = layout_with_kk,  main = paste(i, "KK"))
  
  
  plot(simp_graph, layout = layout_nicely,  main = paste(i, "Nicely"))


}


```


### Other network Visualizations

Here are some other visualizations of the subgraphs displaying certain attributes

```{r}
set.seed(31126)
plot(
  subg1,
  layout = layout_nicely,
  vertex.size=2*log(V(subg1)$target, base=10),
  vertex.label.cex=.4,
  edge.width=2,
  main="subg1 where size is log internet traffic"
)
set.seed(31126)
plot(
  subg1,
  layout = layout_nicely,
  vertex.size=2*log(lengths(V(subg1)$features)),
  vertex.label.cex=.4,
  edge.width=2,
  main="subg1 where size is log length of the word list"
)

```



## Community Detection

[back to report](#community-detection)

```{r}
subg <- list(subg1 = subg1, 
             subg2 = subg2,
             subg3 = subg3,
             subg4 = subg4)

und_subg <- lapply(subg, as.undirected)
```

### We choose Fast Greedy, Louvain, and Walktrap (top three methods on our lecture slide and covered in ica) to perform our community detection.

[back to report](#run-community-detection-algorithms)

```{r}
und_subg <- lapply(subg, function(x) simplify(as_undirected(x)))
```

```{r}
com_fg <- lapply(und_subg, cluster_fast_greedy)
com_l <- lapply(und_subg, cluster_louvain)
com_w <- lapply(und_subg, cluster_walktrap)
```

### And then we compare modularity

[back to report](#analyze-and-interpret-the-resulting-communities)

```{r}
mod_fg <- sapply(com_fg, modularity)
mod_l <- sapply(com_l, modularity)
mod_w <- sapply(com_w, modularity)

data.frame(
subgraph = names(subg),
fast_greedy = mod_fg, 
louvain = mod_l,
walktrap = mod_w
)

ncom_fg <- sapply(com_fg, length)
ncom_l <- sapply(com_l, length)
ncom_w <- sapply(com_w, length)

data.frame(
subgraph = names(subg),
fast_greedy = ncom_fg, 
louvain = ncom_l,
walktrap = ncom_w
)
```

### And as it was recommended in our lecture to report community size, we do so as well here.

[back to report](#analyze-and-interpret-the-resulting-communities)

```{r}
sizes(com_l[[1]])
sizes(com_l[[2]])
sizes(com_l[[3]])
sizes(com_l[[4]])
```

### And we visualize with community coloring, short eg

[back to report](#analyze-and-interpret-the-resulting-communities)

```{r}
plot(com_l[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Louvain community in subgraph 1")
plot(com_l[[2]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Louvain community in subgraph 2")
plot(com_l[[3]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Louvain community in subgraph 3")

plot(com_fg[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Fast Greedy community in subgraph 1")
plot(com_fg[[1]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Fast Greedy community in subgraph 2")
plot(com_fg[[1]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Fast Greedy community in subgraph 3")

plot(com_w[[1]], und_subg[[1]], layout = layout_with_fr(und_subg[[1]]), vertex.label = NA, main = "Walktrap community in subgraph 1")
plot(com_w[[2]], und_subg[[2]], layout = layout_with_fr(und_subg[[2]]), vertex.label = NA, main = "Walktrap community in subgraph 2")
plot(com_w[[3]], und_subg[[3]], layout = layout_with_fr(und_subg[[3]]), vertex.label = NA, main = "Walktrap community in subgraph 3")
```


## Adjacency Matrix

[back to report](#adjacency-matrix)

### Unordered Adjacency Matrix

```{r}
library(Matrix)

non_ord <- com_l[[1]]$membership
A <- as_adj(und_subg[[1]])
image(Matrix(A))
```

### Ordered Adjacency Matrix

```{r}
#library(Matrix)

ord <- order(com_l[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]))

ord <- order(com_fg[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]), main="subg1 adj matrix, fast greedy communities")

ord <- order(com_w[[1]]$membership)
A <- as_adj(und_subg[[1]])
image(Matrix(A[ord, ord]), main="subg1 adj matrix, walktrap communities")
```

## Analyzing Network Metrics (Degree)

[back to report](#degree)

```{r}
hist(degree(g))
```
```{r}
# Degree of Nodes in Subgraph 1
degree <- degree(subg1, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg1)$color <- "maroon"
V(subg1)$frame.color <- "white"


# Highlights top 5 node in terms of degree
V(subg1)$color[V(subg1)[head(degree_sort$ix)]] <- "#F0DC16"


plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```
```{r}
# Degree of Nodes in Subgraph 2
degree <- degree(subg2, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg2)$color <- "maroon"
V(subg2)$frame.color <- "white"

# Highlights top 5 node in terms of degree
V(subg2)$color[V(subg2)[head(degree_sort$ix)]] <- "#F0DC16"


plot(subg2,
     layout = layout_with_kk(subg1),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 3
degree <- degree(subg3, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg3)$color <- "maroon"
V(subg3)$frame.color <- "white"


# Highlights top 5 node in terms of degree
V(subg3)$color[V(subg3)[head(degree_sort$ix)]] <- "#F0DC16"

plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

```{r}
# Degree of Nodes in Subgraph 4
degree <- degree(subg4, mode="all")

# Degree Histogram
hist(degree,
     breaks=50, 
     main="Degree histogram") 

# Network Plot
degree_sort <- sort(degree, decreasing = T, index.return = T)
head(degree_sort$ix)

V(subg4)$color <- "maroon"
V(subg4)$frame.color <- "white"


# Highlights top 5 node in terms of degree
V(subg4)$color[V(subg4)[head(degree_sort$ix)]] <- "#F0DC16"

plot(subg4,
     layout = layout_with_kk(subg4),
     vertex.size = log(degree + 1) * 3,
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```


## Analyzing Network Metrics (EigenCentrality)

[back to report](#eigen-metric)

```{r EigenCentrality, cache=TRUE}
print("Eigen Centrality SubGraph 1", quote=FALSE)
eig_subg1 <- eigen_centrality(subg1)$vector
hist(eig_subg1)
hist(15*sqrt(sqrt(eig_subg1)))
plot(subg1,
     layout = layout_with_kk(subg1),
     vertex.size = 15*sqrt(sqrt(eig_subg1)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 2", quote=FALSE)
eig_subg2 <- eigen_centrality(subg2)$vector
hist(eig_subg2)
plot(subg2,
     layout = layout_with_kk(subg2),
     vertex.size = 15*sqrt(sqrt(eig_subg2)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 3", quote=FALSE)
eig_subg3 <- eigen_centrality(subg3)$vector
hist(eig_subg3)
plot(subg3,
     layout = layout_with_kk(subg3),
     vertex.size = 15*sqrt(sqrt(eig_subg3)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)

print("Eigen Centrality SubGraph 4", quote=FALSE)
eig_subg4 <- eigen_centrality(subg4)$vector
hist(eig_subg4)
plot(subg4,
     layout = layout_with_kk(subg4),
     vertex.size = 15*sqrt(sqrt(eig_subg4)),
     vertex.label=NA,
      edge.arrow.size = 0.1,
     edge.color = "gray80",
     edge.width = 0.5)
```

## Monte Carlo Modularity

[back to report](#monte-carlo-estimate-of-modularity)

```{r, eval=FALSE}
set.seed(123) # For reproducibility
n_sim <- 1000
null_mods <- numeric(n_sim)

for(i in 1:n_sim) {
  # Create a randomized graph preserving degree distribution
  # "degseq" method keeps the degree of each node the same
  rg <- sample_degseq(degree(subg1), method = "vl")
  
  # Run the SAME community detection algorithm on the random graph
  r_comm <- cluster_louvain(rg)
  null_mods[i] <- modularity(r_comm)
}

save(null_mods, file="null_mods.rda")
```

```{r}
load('null_mods.rda')

hist(null_mods)
```





