---
title: "PSTAT 194 Individual Final Messing Around"
author: "Ayden Kelleher"
date: "2026-03-04"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
    highlight: tango
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load and Prepare the Data

## Load your chosen dataset into the course RStudio server and convert it to an ‘igraph’ object.

We loaded the data in [Code Snippet 1](#snippet-1)


## Consider what the vertices and edges represent in your network.

In our network, the vertices represent 


## Identify basic network elements: vertices, edges, weights, directionality, etc

# Generate Induced Subgraphs:

## Create four induced subgraphs with 200 random nodes each from your dataset.

# Visualize the Networks:

## Use three different layouts from the igraph package to visualize each network.

## Experiment with node and edge attributes to enhance visualizations.

# Analyze Network Metrics:

## Explore various network, node, and edge metrics.

## Interpret these metrics in the context of your network.

# Community Detection:

## Run community detection algorithms.

## Analyze and interpret the resulting communities

# Adjacency Matrix:

## Examine the adjacency matrix of your network.

## Reorder vertices to highlight patterns.

# Reorder Nodes:

## Reorder nodes based on community or connected component.

## Observe changes in the adjacency matrix and identify clearer patterns

# Summarize Findings:

## Prepare a presentation and a short write-up summarizing your analysis and findings.

# Appendix

## Snippet 1

```{r}
#install.packages('igraph')
library(igraph)

edges_matrix <- as.matrix(read.csv("musae_squirrel_edges.csv"))

edges_matrix[,1] <- factor(edges_matrix[,1]) # if we dont change this to factor, then there are too many id names idk. If you don't believe me comment this out :/

g <- graph_from_edgelist(edges_matrix, directed = FALSE)

```

```{r}
length(V(g)) #this is exactly what the help page says
length(E(g)) #this is quite a few more edges than the help page says
```

## First Attempt at creating a subgraph

```{r}
set.seed(30426)
subgraph_attempt <- induced_subgraph(g, sample(V(g), 200))

plot(subgraph_attempt,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


## Subgraph 1




```{r}
set.seed(30326)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg1 <- induced_subgraph(g, sampled_vertices[1:200])
```


## Plot 1

```{r}
plot(subg1,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

## Subgraph 2

```{r}
set.seed(30426)
centers <- sample(V(g), 40)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg2 <- induced_subgraph(g, sampled_vertices)
```

## Plot 2

```{r}
plot(subg2,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


## Subgraph 3

```{r}
set.seed(30526)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg3 <- induced_subgraph(g, sampled_vertices)
```

## Plot 3

```{r}

plot(subg3,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```


## Subgraph 4

```{r}
set.seed(30626)
centers <- sample(V(g), 41)
sampled_vertices <- c()

for (v in centers) {
  
  neigh <- neighbors(g, v)
  
  if (length(neigh) > 5) {
    neigh <- sample(neigh, 5)
  }
  
  sampled_vertices <- c(sampled_vertices, v, neigh)
}

subg4 <- induced_subgraph(g, sampled_vertices)
```

## Plot 4

```{r}
plot(subg4,
  vertex.size=10,
  vertex.label.cex=.4,
  edge.width=2)
```

